Functory: Une bibliothèque de calcul distribué \\ pour Objective Caml\footnotetext{Ce travail a été réalisé en partie dans le cadre du projet U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2010-10-13 doc/jfla-2011/main.tex
*** empty log message ***

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

@@

\newcommand{\Ocaml}{OCaml}
\newcommand{\functory}{\textsf{Functory}}
\newcommand{\JoCaml}{Jo{\&\!}Caml}
\newcommand{\unix}{\textsc{Unix}}

\title{Functory: Distributed Computing \\ for the Common Man\footnotetext{Ce
    travail a été réalisé en partie dans le cadre du projet U3CAT
    (\emph{Unification of Critical C Code Analysis Techniques},
    ANR-08-SEGI-021).}}

@@

\begin{abstract}
  We present Functory, a distributed computing library for
  functional programming languages. The main features of this library
  include (1) a polymorphic API, (2) several implementations to
  adapt to different computing environments such as sequential,
  multi-core or network, and (3) a reliable fault-tolerance mechanism.
  This paper describes the motivation behind this work, as well as
  the design and implementation of the library. It also demonstrates
  the potential of the library using realistic experiments.
\end{abstract}

@@

This paper introduces a generic library for distributed computing,
targeted at functional programming. 
This work was initially motivated by the computing needs that exist in our
own research team. Our applications include large-scale deductive
program verification, which amounts to checking the validity of a
large number of logical formulas using a variety of automated theorem
provers~\cite{filliatre07cav}. Our computing infrastructure consists
of a few powerful multi-core machines (typically 8 to 16 cores) and
several desktop PCs (typically dual-core). However, for our
application needs, there is no library that helps in exploiting such a
computing infrastructure in our favorite functional programming
language.  Hence we designed and implemented such a
library, which is the subject of this paper.
This library is implemented in \Ocaml, but the
implementation should be straightforward in any functional
programming language.

The distributed computing library presented in this paper is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of APIs that allows the
execution of large-scale parallelizable computations, very relevant to
our application needs (and presumably relevant to a variety of
real-world applications), over multiple cores in the same machine or
over a network of machines. 
The most important features of our library are the following:
\begin{itemize}
\item \emph{Genericity}: 
  it allows various patterns of polymorphic computations;
\item \emph{Simplicity}: switching between multiple cores on the same
  machine and a network of machines is as simple as changing a couple
  of lines of code;
\item \emph{Task distribution and fault-tolerance}: 
  it relieves the user from implementing task distribution routines
  and reliable fault-tolerance mechanisms.
\end{itemize}
The application domain of such a distributed computing library is manyfold. 
It is worth noting that the library is not targeted at applications
running on server farms, crunching enormous amounts of data. 
It is rather more relevant for research labs, to make use of available
computing resources efficiently. It serves a variety of users and a wide
spectrum of needs, from desktop PCs to networks of machines, and hence
the title.
The most striking application would involve efficient execution
of any large-scale expensive computation that is parallelizable.
This is the case in our
research endeavours, that is validating thousands of verification
conditions using automated theorem provers.

% what it is and what it is not
% \begin{itemize}
% \item it is not a library which does parallelization
% \item it is a library for reliable distributed execution of
%   parallelizable computation 
% \end{itemize}
% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
% programming, ironically) ;  however, there are differences:
% \begin{itemize}
% \item polymorphic, higher-order API
% \item more generic (cores/network)
% \item does not focus on association lists
% \item we don't have data locality (future work)
% \end{itemize}
% what is the target audience/applications: 
% for instance, use if automatic provers on thousands of verification
% conditions, on computing infrastructures which can be
% \begin{itemize}
% \item a single machine with multiple cores, possibly remote
% \item several machines, small or large in number, over a network
% \end{itemize}

@@

% LocalWords:  parallelize functor parameterized indices endianness monomorphic
% LocalWords:  genericity executables OCamlMPI MapReduce

>>>>

@@

\newcommand{\Ocaml}{OCaml}
\newcommand{\functory}{\textsf{Functory}}
\newcommand{\JoCaml}{Jo{\&\!}Caml}
\newcommand{\unix}{\textsc{Unix}}
\newcommand{\coeur}{c\oe ur}
\newcommand{\coeurs}{c\oe urs}

\title{Functory: Une bibliothèque de calcul distribué \\ pour
  Objective Caml\footnotetext{Ce
    travail a été réalisé en partie dans le cadre du projet U3CAT
    (\emph{Unification of Critical C Code Analysis Techniques},
    ANR-08-SEGI-021).}}

@@

\begin{abstract}
  Cet article présente \functory, une bibliothèque de calcul distribué
  pour Objective Caml. Les principales caractéristiques de cette
  bibliothèque sont (1) une interface polymorphe, (2) plusieurs
  réalisations correspondant à des contextes d'utilisation différents
  et (3) un mécanisme de tolérance aux pannes.
  Cet article détaille la conception et la réalisation de Functory et
  montre son potentiel sur de nombreux exemples.
\end{abstract}

@@

Cet article présente \functory, une bibliothèque de calcul distribué
pour Objective Caml. Initialement, ce travail a été motivé par des
besoins de calcul au sein de notre équipe de recherche, ProVal.
Nos applications en vérification déductive de programmes incluent
notamment la validation de très nombreuses formules logiques par des
démonstrateurs automatiques variés~\cite{filliatre07cav}.
Nos moyens de calcul consistent en quelques machines très puissantes
(typiquement 8 ou 16 \coeurs) et plusieurs ordinateurs de bureau
(typiquement 2 \coeurs). Aucune bibliothèque ne nous permettait de
tirer facilement partie de cette infrastructure de calcul dans notre
langage préféré. C'est pourquoi nous avons conçu et réalisé la
bibliothèque \functory\ qui est le sujet de cet article.
Cette bibliothèque est réalisée pour \Ocaml\ mais pourrait facilement
être adaptée pour tout autre langage fonctionnel.

\functory\ n'est pas une bibliothèque qui aide l'utilisateur à
paralléliser ses calculs. Son rôle consiste plutôt à offrir des
facilités pour distribuer, de manière sûre, des calculs déjà
identifiés comme indépendants. En particulier, \functory\ offre
plusieurs interfaces génériques pour distribuer des calculs sur différents
\coeurs\ d'une même machine ou sur un réseau de machines. Ceci
correspond exactement au contexte qui a motivé la construction de
cette bibliothèque mais aussi, très probablement, à celui de
nombreuses autres applications.
Les principales caractéristiques de \functory\ sont les suivantes :
\begin{itemize}
\item \emph{généricité} : les
  traits fonctionnels que sont l'ordre
  supérieur et le polymorphisme sont exploités pour fournir un maximum
  de généricité ;
\item \emph{simplicité} : on passe d'un calcul séquentiel à un
  calcul multi-\coeurs puis à un calcul en réseau en ne modifiant
  que quelques lignes dans le code ;
\item \emph{distribution et tolérance aux pannes} : l'intégralité de
  la gestion de la distribution et de la tolérance aux pannes est
  prise en charge par la bibliothèque.
\end{itemize}

The application domain of such a distributed computing library is manyfold. 
It is worth noting that the library is not targeted at applications
running on server farms, crunching enormous amounts of data. 
It is rather more relevant for research labs, to make use of available
computing resources efficiently. It serves a variety of users and a wide
spectrum of needs, from desktop PCs to networks of machines, and hence
the title.
The most striking application would involve efficient execution
of any large-scale expensive computation that is parallelizable.
This is the case in our
research endeavours, that is validating thousands of verification
conditions using automated theorem provers.

@@

% LocalWords:  parallelize functor parameterized indices endianness monomorphic
% LocalWords:  genericity executables OCamlMPI MapReduce généricité

%%% Local Variables: 
%%% mode: latex
%%% ispell-local-dictionary: "francais"
%%% TeX-master: t
%%% End:
